---
title: "2018 Mass Shootings"
author: "Arjun Vaghela"
date: "05/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
# Source in the other script
source("analysis.R")

# Loading `knitr` library
library(knitr)

```

## Summary

In 2018, the United States had a large number of shootings that caused many
injuries and deaths. There were a total of `r total_shootings` shootings that
occurred during this year and a total of `r total_lives_lost` people lost their
lives. With numerous shootings happening during the year, 
`r most_shooting_occurrences` seemed to be impacted the most as they had the 
most occurrences out of any other city. Within Chicago, there were a total of 
`r chicago_injured` injured due to the shootings along with `r chicago_killed`
killed during 2018.

## Summary Table
With many cities and states finding themselves having to recover from these
shootings, below you can find an aggregated table that shows which states 
have suffered the most in terms of people killed and injured in 2018.


```{r, echo=FALSE}
kable(summary_table)
```

From this table, you can see how different states were impacted based on the
on the total amount of injured and killed people. 

## Parkland Shooting
On `r parkland_date`, Nikolas Cruz opened fire at Marjory Stoneman Douglas 
High School in `r parkland_city`, `r parkland_state`. Reports say that the 
coordinate points for the shooting were around 
(`r parkland_lat`, `r parkland_long`), which confirms that it was indeed on the 
school premises. This tragedy ended up impacting many people and in total, Cruz
ended up killing `r parkland_killed` and injuring `r parkland_injured` students
and staff. According to NBC News, Cruz had reportedly been expelled from the 
high school and was enrolled at another school at the time of the shooting. 
This shooting was was arguably the largest shooting in history and even 
President Donald Trump tweeted that there were "many signs the Florida
shooter was mentally disturbed" (https://www.nbcnews.com/news/us-news/police-respond-shooting-parkland-florida-high-school-n848101).

## Interactive Map

```{r, echo=FALSE}
interactive_map
```

Above you will find an interactive map that shows the shootings that had 
occurred during 2018. As you hover over the different markers, you will find
information about each shooting. The Parkland shooting described earlier
is shown with the largest circle on the map as it was the largest shooting in
2018, followed by other large circles that represent other large shootings. 

## States with Highest Shootings (Bar Graph)
With many students going out of state for college, safety is one of the main 
concerns that people take into account. Many people wonder which states are the
most prone to gun violence and below you will find a graph that includes the 
top 6 states with the highest amount of deaths due to shootings in 2018. 

```{r, echo=FALSE}
chart
```

As shown in the chart above, California and Florida were the two highest ranked 
states in terms of deaths and injured people due to gun violence. With
California having 50 deaths and Florida have 48 deaths in 2018, we can infer
that these two states have the highest chance of people being killed in 
shootings. We can also see that within the top 6 highest death states, 
Tennessee (19 deaths) and Illinois (21 deaths) had a large difference in deaths
related to shootings compared to California and Illinois.

